Mapping factors associated with adolescents' emotion dysregulation: A text-mining based systematic review

### 2020-06-07 manuscript.tex
Final screening done

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Mapping the predictors of adolescent emotion dysregulation: A text-mining based systematic review},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{Mapping the predictors of adolescent emotion dysregulation: A text-mining based systematic review}

@@

A key developmental challenge in adolescence is acquiring mature emotion regulation skills (Crone \& Dahl, 2012; Zimmermann \& Iwanski, 2014). Emotion regulation is integral to mental health(Aldao, Nolen-Hoeksema, \& Schweizer, 2010; Braet et al., 2014; Schäfer, Naumann, Holmes, Tuschen-Caffier, \& Samson, 2017) and social functioning (Reindl, Gniewosz, \& Reinders, 2016). For as many as one in five adolescents, this phase marks the onset of emotion regulation-related mental illness, which can persist throughout the life course (see Lee et al., 2014). It is therefore crucial to identify which risk factors and environmental hazards render some adolescents more susceptible to emotional difficulties than others. Although there is an abundance of empirical work on factors associated with adolescents' emotion regulation, the literature is somewhat fragmented, as the topic has been investigated from different (sub)disciplines. This study seeks to address this limitation by mapping all factors currently thought to be relevant to adolescent emotion regulation, based on a systematic review of the literature. Similarly, although there are several \emph{relevant} theories, there is no specific theoretical framework to guide research on adolescents' emotion regulation. Although theory development is beyond the scope of this paper, we set out to generate a proto-theoretical nomological network of the potential risk factors, manifestations, and outcomes of emotion regulation in adolescence. We discuss how this nomological network relates to existing relevant theory, and conclude with recommendations for future theory-generating efforts.

There have been only few notable efforts to provide an encompassing framework (Buss, Cole, \& Zhou, 2019; Coe-Odess, Narr, \& Allen, 2019; Riediger \& Klipker, 2014). Yet several limitations remain.

instead takes an inductive approach to map the correlates of adolescents' emotion regulation based on all available literature.

The present study instead takes an inductive approach to map the correlates of adolescents' emotion regulation based on all available literature. From this mapping, we proceed to classify correlates as potential risk factors, manifestations, and outcomes of adolescent emotion regulation. The results provide a conceptual overview of the existing literature, can help identify blind spots in existing theory, and might inspire new hypotheses to guide future deductive research.

Several limitations emerge from prior efforts to provide an encompassing framework of this literature

These efforts at unification have, thusfar, been conducted in a top-down, theory-driven manner.

@@

Several recent publications have undertaken the monumental task of providing a comprehensive overview of theories of emotional development (e.g., Buss et al., 2019; {\textbf{???}}; Riediger \& Klipker, 2014), and of the empirical work pertaining to emotions in adolescence (Coe-Odess et al., 2019). We provide a summary review of the theoretical landscape as context for our empirical contribution.

Two of the most frequently cited theories in research on development, including emotional development, are biopsychosocial models. One is Bronfenbrenner's bioecological model, which describes how the environment shapes individual development. The individual is imbued with certain biological predispositions, and changes over the course of developmental time. Moreover, this individual development takes place in interaction with contextual influences. These range from the microsystem, composed of other people in close interaction with the individual, to the macrosystem, consisting of indirect political and economic influences, to the exosystem, consisting of cultural norms and values. Sameroff's transactional model is another bioecological theory, which focuses heavily on social influences (the microsystem in Bronfenbrenner's model). Its main contribution is that it conceptualized development as a product of reciprocal influences between child and parent (or another individual).

With regard to theories specifically addressing emotion regulation in adolescence, Hall's notion of \enquote{storm and stress} is one of the oldest. It describes how hormonal changes diminish self-control and increase reactivity, leading to difficulties in emotion regulation, conflict with parents, and risky behavior. Recent work provides a more nuanced perspective on this theory. For example, although the notion of diminished self-control and increased emotional reactivity has been upheld, it has been recast as a necessary change that facilitates emotional maturation (Crone \& Dahl, 2012), at the risk of emotional disturbance (Arnett, 1999; Lee et al., 2014). This again highlights the importance of identifying factors that predict or explain between-adolescent differences in emotional dysregulation.

Another theory of emotion regulation development was proposed by Sroufe (1995). This theory focuses on normative emotional development in early childhood. It argues that children's self-regulatory abilities increase with age, which drives a transition from external emotion regulation by primary caregivers towards autonomous emotion regulation. This work describes social and cognitive influences. Social influences mainly occur through parents' co-regulation, parenting behaviors, and parent-child attachment. Cognitive influences occur through the development of the central nervous system (CNS), cognition, and self-regulation. This emphasis on neorocognitive development is mirrored in polyvagal theory (Porges, 1995), which closely links emotional experience - and regulation - to autonomous nervous system functioning.

A more recent neorocognitive theory, with greater relevance to adolescents' emotion regulation development, is Crone and Dahl's model of social-affective engagement and goal flexibility (Crone \& Dahl, 2012). This theory is based on the notion of a \enquote{maturity gap} in middle-adolescence, arising from a developmental asymmetry between motivational and inhibitory brain circuits ({\textbf{???}}; Cracco, Goossens, \& Braet, 2017). However, it explicitly sets out to explain adolescents' diverging destinies - the observation that some youngsters flourish while others languish. Crone and Dahl argue that adolescents' cognitive engagement is dynamically responsive to social and motivational goal salience. This flexibility, on the one hand, prepares adolescents to effectively engage cognitive systems in novel challenging situations in a way that facilitates developing mature regulatory abilities. On the other hand, it places them at risk to act impulsively in pursuit of peer approval. This theory focuses primarily on cognitive factors and the role of peers, with little reference to parenting or other factors.

@@

Holodynski and Friedlmeier's ({\textbf{???}}) internalization theory applies Vygotsky's theory of development to the domain of emotion. Like Sroufe's work, it captures the transition from interpersonal to intrapersonal emotion regulation. Unique features of this work include, most notably, the integrative view of emotion and emotion regulation, the interplay between emotion and communication, and the role of culture in emotion regulation development.

Aside from theories of emotional development, theories of the phenomenon of emotion regulation are also somewhat relevant. One influential theory is Gross' (2013) process model, which describes the process of emotion regulation, from eliciting cue to ultimate response. Individuals modulate the different stages of this process using strategies, consciously or otherwise. The effectiveness, desireability, and consequences of different strategies depend on the cultural context (see Bariola, Gullone, \& Hughes, 2011). Similar to Gross' theory, the social information processing theory of emotion also describe the role of intrapsychic processes and strategies in emotion experience and reulation (Lemerise \& Arsenio, 2000). Such theories are particularly relevant to a fine-grained understanding of the process of emotion regulation, but their developmental relevance has unfortunately not yet been considered (REF Buss).

Parents are widely considered to be the primary proximal influence driving emotion regulation development.

\hypertarget{shortcomings}{%
\subsection{Shortcomings}\label{shortcomings}}

Extant theoretical models have several noteable shortcomings. The most notable shortcoming is the lack of explicit attention to the life phase of adolescence. This is unfortunate, as empirical research consistently indicates that adolescence is a crucial life phase for emotion regulation development. As this life phase differs qualitatively from infancy, childhood, and adulthood, it is insufficient to simply extrapolate from theories focused on these ages. Other theories do focus on adolescence specifically, but only address emotion regulation in passing. A second, related, shortcoming is that some relevant theories lack a developmental component, or are based on a near-exclusively cross-sectional body of research (see Buss et al., 2019; Crone \& Dahl, 2012). A third shortcoming is that theories vary widely in scope: On the macro-level, there are theories such as the biopsychosocial models of Bronfenbrenner and Sameroff, which are all-encompassing but non-specific. On the micro-level, there are theories such as Morriss' tripartite model, Crone and Dahl's model, and polyvagal theory, which originated in subdisciplines and describe one facet of the issue in great detail, but have not been placed within a unifying framework (Buss et al., 2019). To conclude, one limitation echoed across many publications in this field is that more theory formation is required to provide a unified framework that could guide future empirical work. A systematic conceptual review of the field could address this shortcoming.

\hypertarget{relevant-factors}{%
\subsection{Relevant factors}\label{relevant-factors}}

As there are few theories specifically relevant to the field of adolescent emotion regulation development, it is perhaps unsurprising that some literature has proceeded in a somewhat less theoretical manner. Such studies do often focus on specific factors that are considered relevant for emotion regulation development.

These include neurological (Fox 1994;
Porges et al.~1994; Quirk 2007; Stansbury and Gunnar
1994), genetic (Goldsmith et al.~1997; Hariri and Forbes
2007), and temperamental inﬂuences (Calkins 2004)
A range of social factors have been proposed to be linked to
ER development, including interactions with parents,
teachers, and peers, as well as more distal societal inﬂu-
ences such as culture and the media (Eisenberg and Morris
2002; Klimes-Dougan et al.~2007; Morris et al.~2007;
Thompson 1994).

on adolescent emotion regulation development is limited27--29. The field is therefore at an impasse: We know that some adolescents are more susceptible to emotion dysregulation than others, but lack tools and theory to identify important predictors of individual development30.

\hypertarget{the-present-study}{%
\subsection{The present study}\label{the-present-study}}

The present paper aims to develop a nomological network of the constructs related to adolescent emotion dysregulation by mapping the available literature. A nomological network is a diagrammatic representation of a phenomenon; a proto-theoretical device that describes relationships between constructs relevant to the theory (REF: Cronbach \& Meehl 1955; McKenna, 2006; Alavi 2018). It can be useful as a precursor to a more explicit theory of the phenomenon. The present paper uses an inductive approach to develop this nomological network. First, we conduct a systematic search to elicit a corpus of relevant literature. Second, we extract constructs from the corpus. Third, we use a dictionary to pare down the extracted constructs to meaningful superordinate categories. Fourth, we map interrelations between these constructs. Fifth, we classify contstructs as potential predictors or outcomes, and we categorize predictors by level of analysis (CJ: Sameroff, which level of biopsychosocial).

@@

This string returned 6653 results, and matched 25 of the 29 records in the reference set. Three search terms could be added to match all 29 reference set items. The terms \texttt{"emotio*\ socialization"\ OR\ "emotio*\ processes"}, as synonyms of emotion regulation, added 191 new hits. However, many of these hits were not directly relevant to emotion regulation. The term \texttt{"development*}, as synonym of adolescence, added 3628 new hits, many of which were not focused on the age range of adolescence. We thus deemed these terms to be overly inclusive, and proceeded with the original search string above.

@@

Duplicates were detected based on exact DOI matches (2 duplicates), and title similarity (54 duplicates). Manual screening in Rayyan resulted in the removal of an additional 13 duplicates.

@@

\begin{verbatim}
## [1] "2000-8198" "1540-2002" "2152-0828" "2152-0828" "0021-5368" "0021-9630"
\end{verbatim}

We manually screened 500 abstracts for suitability in Rayyan. Then, we used ASReview to rank the predicted relevance of all papers based on the manually included and excluded articles. We screened the lowest-ranked articles to find an inflection point below which most articles were irrelevant, and excluded those from analysis.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-aldaoEmotionregulationStrategiesPsychopathology2010}{}%
Aldao, A., Nolen-Hoeksema, S., \& Schweizer, S. (2010). Emotion-regulation strategies across psychopathology: A meta-analytic review. \emph{Clinical Psychology Review}, \emph{30}(2), 217--237.

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Mapping factors associated with adolescents' emotion dysregulation: A text-mining based systematic review},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{Mapping factors associated with adolescents' emotion dysregulation: A text-mining based systematic review}

@@

A key developmental challenge in adolescence is acquiring mature emotion
regulation skills (Crone \& Dahl, 2012; Zimmermann \& Iwanski, 2014). Emotion regulation is integral to
mental health(Aldao, Nolen-Hoeksema, \& Schweizer, 2010; Braet et al., 2014; Schäfer, Naumann, Holmes, Tuschen-Caffier, \& Samson, 2017)
and social functioning (Reindl, Gniewosz, \& Reinders, 2016). For as many
as one in five people, adolescence marks the onset of emotion
regulation-related mental illness, which can persist throughout the life course
(see Lee et al., 2014). It is therefore crucial to identify which
risk factors and environmental hazards render some adolescents more susceptible
to emotional difficulties than others. Although there is an abundance of
empirical work on adolescents' emotion regulation, the literature is somewhat
fragmented, as the topic has been investigated from different (sub)disciplines.
This study seeks to overcome this limitation by mapping all factors currently
thought to be relevant to adolescent emotion regulation, based on a systematic
review and bibliographic analysis.

\hypertarget{why-adolescence-is-an-important-life-phase-for-emotion-regulation}{%
\subsection{Why adolescence is an important life phase for emotion regulation}\label{why-adolescence-is-an-important-life-phase-for-emotion-regulation}}

Adolescence is defined as a life phase ranging from pubertal onset to
adult-like independence (Steinberg, 2014). This definition
invokes both biological- and socio-cultural factors. It has been argued that,
due to accellerated pubertal onset and delayed transitions to adult roles,
adolescence now ranges from 10-24 years in the modern Western world
(Sawyer, Azzopardi, Wickremarathne, \& Patton, 2018).

Adolescence is increasingly seen as a developmentally sensitive period for
mature emotion regulation skills. During this period, children experience rapid
biological, cognitive, and social changes that prompt new emotional
experiences, and tax existing regulatory abilities
(Steinberg, 2014). At the same time, the staggered
development of motivational-emotional and regulatory brain circuits gives rise
to a maturity gap, where youngsters pursue new experiences in life and love,
without being fully prepared to cope with the emotional outcomes
(Crone \& Dahl, 2012). Due to these coalescing changes,
adolescents display a restricted repertoire of emotion regulation strategies
compared to younger or older individuals
(Zimmermann \& Iwanski, 2014). Consequently, adolescents experience
more frequent and intense (negative) emotions than children or adults (see Silk, Steinberg, \& Morris, 2003). The literature thus paints a picture of
adolescence as a chrysalis for emotional development: Children enter this stage
with emotion regulation skills adapted to the challenges of childhood. During
adolescence, emotional systems are rearranged substantially. These changes
render adolescents temporarily vulnerable to emotional dysregulation, but
ultimately serve a functional role in acquiring mature emotion regulation
skills. Eventually, most adolescents emerge ready to take on the challenges of
early adulthood. Unfortunately, a sizeable proportion of adolescents instead
manifests enduring emotional difficulties.

\hypertarget{the-consequences-of-difficulties-in-emotion-regulation}{%
\subsection{The consequences of difficulties in emotion regulation}\label{the-consequences-of-difficulties-in-emotion-regulation}}

The
consequences of difficulties in emotion regulation are extensive,
well-documented, and persist throughout the life course. Meta-analyses indicate
that such difficulties are associated with externalizing, and more strongly,
with internalizing psychopathology
(Aldao et al., 2010; Schäfer et al., 2017). Individuals who engage in maladaptive
emotion regulation strategies also experience more negative emotions,
diminished well-being, and greater strain in interpersonal relationships
(Bell \& Calkins, 2000; Gross \& John, 2003).
Given the
prevalence of emotional difficulties and their extensive implications for
individual mental health, wellbeing, and social functioning, and downstream
cost to society as a whole (Lee et al., 2014), it is essential to
have a solid understanding of the risk factors associated with adolescents'
emotion regulation difficulties.

@@

Theory is of paramount importance when seeking to understand or explain a
phenomenon. Although there are many relevant theories, none have specifically
addressed emotion regulation development in adolescence. Several recent
publications have undertaken the monumental task of comprehensively reviewing
theory relevant to emotional development
(Buss, Cole, \& Zhou, 2019; Riediger \& Klipker, 2014). Thanks, in part, to their efforts,
we can provide a summary review of the theoretical landscape as context for our
empirical contribution.
Two theories widely invoked to frame research on development, including
emotional development, are Bronfenbrenner's bioecological model, and Sameroff's
transactional model. Bronfenbrenner's model
(Bronfenbrenner \& Morris, 2007) describes how the environment
shapes individual development. Each individual is imbued with certain
biological predispositions, and develops over time, in interaction with
contextual influences. These influences range from the microsystem, composed of
those people close to the individual, to the macrosystem, consisting of
indirect political and economic influences, to the exosystem, consisting of
cultural norms and values. Sameroff's model is somewhat compatible with the
bioecological model ({\textbf{???}}), but focuses more
on the interplay between nature and nurture. It conceptualizes development as a
product of reciprocal influences between the child and the environment.
Sameroff distinguishes between proximal influences, for instance, through
parents (the microsystem in Bronfenbrenner's model), and distal influences;
structural factors indirectly shaping development, like socio-economic status,
schools, and the community (macro- and exosystem). With increasing age, distal
influences gradually gain ground on proximal influences.

Turning now towards theories specifically addressing emotion regulation in
adolescence, Hall's notion of \enquote{storm and stress} is one of the oldest (see Arnett, 1999). It describes how hormonal changes diminish
self-control and increase reactivity, leading to difficulties in emotion
regulation, conflict with parents, and risky behavior. Recent work provides a
more nuanced perspective on this theory. For example, although the notion of
diminished self-control and increased emotional reactivity has been upheld, it
has been recast as a necessary change that facilitates emotional maturation
(Crone \& Dahl, 2012), at the risk of emotional
disturbance (Arnett, 1999; Lee et al., 2014).
More recently, Sroufe (1995) presented
a theory focused on normative emotional development in early childhood. It
argues that, as children grow older, their increasing self-regulatory abilities
drive a transition from external emotion regulation by primary caregivers
towards autonomous emotion regulation. This theory focuses on two drivers of
development: social and cognitive influences. Social influences mainly occur
through parents' co-regulation, parenting behaviors, and parent-child
attachment. Cognitive influences occur through the development of the central
nervous system (CNS), cognition, and self-regulation. Several other theories,
introduced below, have extemporized on these sources of influence.

The emphasis on neorocognitive development as a driver of emotional
development, including emotion regulation development, is echoed in several
other models. For example, polyvagal theory closely links emotional experience
- and regulation - to autonomous nervous system functioning
(Porges, 1995). A more recent neorocognitive theory, with
substantial relevance to adolescents' emotion regulation development, is Crone
and Dahl's model of social-affective engagement and goal flexibility
(Crone \& Dahl, 2012). This theory is based on the notion
of a \enquote{maturity gap} in middle-adolescence, arising from a developmental
asymmetry between motivational and inhibitory brain circuits
({\textbf{???}}; Cracco, Goossens, \& Braet, 2017). However, it sets out to explain - to
a greater extent than other writings - adolescents' diverging destinies; the
question why some youngsters flourish while others languish. Crone and Dahl
argue that adolescents' cognitive engagement is dynamically responsive to
social and motivational goal salience. This flexibility, on the one hand,
prepares adolescents to effectively engage cognitive systems in novel
challenging situations in a way that facilitates developing mature regulatory
abilities. On the other hand, it places them at risk to act impulsively in
pursuit of peer approval. This theory focuses primarily on cognitive factors
and the role of peers, with less attention to factors such as parenting.

@@

Aside from theories of emotional development, theories of the phenomenon of
emotion regulation are also somewhat relevant. One influential theory is Gross'
(2013) process model, which describes the
process of emotion regulation, from eliciting cue to ultimate response.
Individuals modulate the different stages of this process using strategies,
consciously or otherwise. The effectiveness, desireability, and consequences of
different strategies depend on the cultural context (see Bariola, Gullone, \& Hughes, 2011). Similar to Gross' theory, the social
information processing theory of emotion also describe the role of intrapsychic
processes and strategies in emotion experience and reulation
(Lemerise \& Arsenio, 2000).

Another theory worth mentioning in this context is
Holodynski and Friedlmeier (2006)'s internalization model of emotional
development. It applies Vygotsky's theory of development to the domain of
emotion, and presents a detailed integrated model of emotional experience and
regulation. Like Sroufe's work, this work describes a transition from
interpersonal to intrapersonal emotion regulation. Other important features of
this work include the interplay between emotion and communication, and the
internalization of the socio-cultural symbolic function of emotion throughout
development.

\hypertarget{shortcomings-of-the-theory}{%
\subsection{Shortcomings of the theory}\label{shortcomings-of-the-theory}}

Existing theoretical models have several limitations. One crucial shortcoming
is the lack of explicit attention to the life phase of adolescence. This is
unfortunate, as empirical research consistently indicates that adolescence is a
crucial life phase for emotion regulation development. Adolescence differs
qualitatively from infancy, childhood, and adulthood. It is therefore
insufficient to simply extrapolate theories focused on these ages. Other
theories do focus on adolescence specifically (e.g., Crone \& Dahl, 2012), but only address emotion regulation
in passing. A second, related, shortcoming is that some relevant theories lack
a developmental component (see Buss et al., 2019; Crone \& Dahl, 2012). A third shortcoming is that theories
vary widely in scope: Some are very broad and all-encompassing, but somewhat
non-specific, such as the biopsychosocial models of Bronfenbrenner and
Sameroff. Others describe a specific phenomenon in great detail, but are not
embedded in a larger unifying framework. Examples include Morris' tripartite
model, Crone and Dahl's model, and polyvagal theory. It would be beneficial to
bridge these two levels of theory. One limitation echoed across many
publications in this field is that more theory formation is required to provide
a unified framework that could guide future empirical work. The present study
seeks to aid such theory development, by providing an inductive conceptual
review of the field.

\hypertarget{bottom-up-insights-from-narrative-reviews}{%
\subsection{Bottom-up insights from narrative reviews}\label{bottom-up-insights-from-narrative-reviews}}

While theory provides a top-down understanding of an area of research,
literature reviews can provide a bottom-up, inductive understanding. Given the
noted absence of a single overarching theory, reviews might provide additional
insight into factors considered relevant in the etiology of adolescents'
emotion regulation competencies. For example, a review by
Bariola et al. (2011) addresses individual psychological factors
such as temperament, and biological factors like neurocognitive development and
genetic predispositions. Furthermore, they address proximal social influences,
including parents, teachers, and peers, and specific mechanisms through which
these individuals exert their influence, including parenting and modeling.
Finally, they address distal factors, including the role of culture and the
media. Barriola and colleagues also offer recommendations for future research,
calling for research on the role of parents beyond early
childhood, and including the role of fathers - a limitation which has since
been addressed by a new wave of empirical research on fathers' role in
emotional development (Van Lissa \& Keizer, 2020; Van Lissa, Keizer, Van Lier, Meeus, \& Branje, 2019).

Another review, by Coe-Odess, Narr, and Allen (2019), offers a nuanced
discussion of several issues that complement prior publications in this field.
For instance, the authors cover the implications of physiological changes in
detail, including neurocognitive development and pubertal maturation. As
adolescents become increasingly individuated, conflict with parents peaks. This
conflict, in turn, impacts emotion regulation, both in terms of day-to-day mood
variability and dispositional emotion dysregulation (see also Van Lissa, Hawk, Koot, Branje, \& Meeus, 2017). As parental importance diminishes,
adolescents become increasingly oriented towards peers. This includes increased
sensitivity to social status and norms, along with the associated increase in
peer pressure and risk taking. Pubertal development also initiates the
emergence of sexual and romantic behavior, and the intensification of both
biological sex differences and gender stereotyped behavior. With regard to the
process of emotion regulation, Coe-Odess and colleagues discuss the importance
of strategies, and point out that, in addition to negative emotionality,
positive emotionality also peaks in adolescence. Relatedly, the authors
describe how hormonal changes precipitate an increased stress response, which
might help explain adolescents' greater susceptibility to emotion
dysregulation. Finally, going beyond the implications of cognitive development
discussed in other publications, the authors discuss how an greater capacity
for abstract thought relates to identity formation - a key challenge in
adolescence ({\textbf{???}}) - and to increased emotional
understanding, and by extension, empathy (see also Van Lissa et al., 2014).

There are notable parallels between the relevant factors highlighted by these
narrative reviews, and the aforementioned theoretical literature. Nevertheless,
these literature reviews also touch upon issues that have received little
attention in theoretical work. This highlights the value of both top-down and
bottom-up approaches to understanding the phenomenon of adolescents' emotion
regulation. One remaining limitation is that, to date, all reviews in this
field have been unstructured narrative reviews. The present publication seeks
to further complement preceding work by conducting a comprehensive empirical
literature review, and mapping the factors associated with adolescent emotion
regulation.

\hypertarget{the-present-paper}{%
\subsection{The present paper}\label{the-present-paper}}

The present paper aims to develop a nomological network of the constructs
related to adolescent emotion dysregulation, based on a systematic review of
the literature. A nomological network is a diagrammatic representation of a
phenomenon; a proto-theoretical device that describes relationships between
constructs relevant to the theory (see Alavi, Archibald, McMaster, Lopez, \& Cleary, 2018). It is a precursor to a more
explicit theory of the phenomenon. The present paper uses an inductive approach
to develop this nomological network. First, we conduct a systematic search to
elicit a corpus of relevant literature. Second, we use text mining techniques
to extract relevant constructs from the corpus. Third, we use a dictionary to
pare down the extracted constructs to meaningful superordinate categories.
Fourth, we map interrelations between these constructs. Fifth, we classify
contstructs as potential predictors or outcomes, and we categorize predictors
by level of analysis.

@@

This string returned 6653 results, and matched 25 of the 29 records in the
reference set. Three search terms could be added to match all 29 reference set
items. The terms \texttt{"emotio*\ socialization"\ OR\ "emotio*\ processes"}, as synonyms
of emotion regulation, added 191 new hits. However, many of these hits were not
directly relevant to emotion regulation. The term \texttt{"development*"}, as synonym
of adolescence, added 3628 new hits, many of which were not focused on the age
range of adolescence. We thus deemed these terms to be overly inclusive, and
proceeded with the original search string above.

@@

Duplicates were identified based on exact DOI matches (2
duplicates), and title similarity (54 duplicates).
Manual screening in Rayyan resulted in the removal of an additional 13 duplicates.

@@

Papers were screened based on two main criteria: First, papers had to address emotion regulation, emotional problems, or similar. Second, the target population must overlap (at least in part) with adolescence, defined as the age range from 10-25. Thus, for example, an article reviewing the role of hormonal changes in emotional adjustment throughout the entire life course, including adolescence, would be deemed relevant. Screening was initially conducted in Rayyan QRCI ({\textbf{???}}). Papers were sorted using Rayyan's ranking algorithm. To train the algorithm on exclusion criteria, screening focused on abstracts most likely to be excluded. After coding 559 papers, screening continued in ASReview - a free, open source alternative to Rayyan QRCI ({\textbf{???}}). This program uses text mining and active learning to build a customizable machine learning model for article inclusion. We used a naive Bayes model, and focused on exclusion of irrelevant papers. In ASReview, an additional 541 papers were screened. Screening again focused on articles most likely to be excluded, and continued until, among the most recently screened 100 papers, only 6 were excluded. The screening procedure is detailed in Figure \ref{prismachart}. In total, 6305 papers were retained for analysis.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/prismachart-1.pdf}
\caption{\label{fig:prismachart}Record screening flowchart}
\end{figure}

\hypertarget{analysis-1-keyword-mapping}{%
\section{Analysis 1: Keyword mapping}\label{analysis-1-keyword-mapping}}

For this first analysis, we relied on author-provided keywords.
We extracted keywords by document, and used a dictionary approach
to classify closely related terms.

\hypertarget{analysis-2-abstract-text-mining}{%
\section{Analysis 2: Abstract text mining}\label{analysis-2-abstract-text-mining}}

This second analysis focused on the abstracts of our corpus. Keywords are
carefully chosen to capture the core tenets of a study. However, as authors are
typically limited to 5 keywords, some of a study's nuance may be lost.
Abstracts offer greater freedom of expression, which is a challenge when it
comes to extracting relevant information. We took a two-pronged approach to
extract relevant topic terms from the abstracts. First, we used the natural
language processing technique part-of-speech tagging (POS-tagging) to identify
nouns in the abstract. Using nouns only in text mining can lead to more
interpretable models (Martin \& Johnson, 2015). Secondly, we used
stemming to reduce these nouns to their root form. Finally, we applied a
dictionary approach as before.

offer an incomplete understanding of a study topic due to the limite
authors are typically limited to provide a maximum of
abstracts offer a little more

used part-of-speech tagging, a natural language processing technique

to identify nouns in the abstracts.
, we relied on author-provided keywords. We extracted keywords by document, and used a dictionary approach to classify closely related terms.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-alaviAligningTheoryMethodology2018}{}%
Alavi, M., Archibald, M., McMaster, R., Lopez, V., \& Cleary, M. (2018). Aligning theory and methodology in mixed methods research: Before Design Theoretical Placement. \emph{International Journal of Social Research Methodology}, \emph{21}(5), 527--540. \url{https://doi.org/10.1080/13645579.2018.1435016}

\leavevmode\hypertarget{ref-aldaoEmotionregulationStrategiesPsychopathology2010}{}%
Aldao, A., Nolen-Hoeksema, S., \& Schweizer, S. (2010). Emotion-regulation strategies across psychopathology: A meta-analytic review. \emph{Clinical Psychology Review}, \emph{30}(2), 217--237.

@@

\leavevmode\hypertarget{ref-bellRelationshipsInputsOutputs2000}{}%
Bell, K. L., \& Calkins, S. D. (2000). Relationships as Inputs and Outputs of Emotion Regulation. \emph{Psychological Inquiry}, \emph{11}(3), 160--163.

@@

\leavevmode\hypertarget{ref-bronfenbrennerBioecologicalModelHuman2007}{}%
Bronfenbrenner, U., \& Morris, P. A. (2007). The bioecological model of human development. In \emph{Handbook of Child Psychology}. John Wiley \& Sons, Inc. \url{https://doi.org/10.1002/9780470147658.chpsy0114}
